---
title: "Reproducible Research: Peer Assessment 1"
author: "Debdulal Dey"
date: "6/22/2019"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}

## Download File
zipurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("./project")){
  dir.create("./project")
}
destFile <- "./project/factivity.zip"

download.file(zipurl,destFile)
unzip(zipfile=destFile,exdir="./project")

## Load Data
sourceFile = file.path("./project", "activity.csv")
list.files("./project")
activity <- read.csv(sourceFile)
summary(activity)
head(activity)
```

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.

1: Make a histogram of the total number of steps taken each day

2: Calculate and report the mean and median total number of steps taken per day

```{r}
library(dplyr)

steps_day<-
  activity%>%
  group_by(date) %>%
  summarize(total_steps = sum(steps)
  ) 
head(steps_day)
```

### Histogram [Total Steps Par Day]

```{r}
library(ggplot2)
hist(steps_day$total_steps, col = steps_day$date,main="Total Steps Per Day", xlab="Total Steps")
```
  
## mean
```{r}
mean(steps_day$total_steps)
```
## median
```{r}
median(steps_day$total_steps)
```

## What is the average daily activity pattern?

1: Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
# aggrgate average data over interval
avg_steps_interval <- aggregate(steps ~ interval, activity, mean)
qplot(interval, steps, data = avg_steps_interval,geom=c("line")) + geom_smooth() + labs(title = "Average Daily Steps") + labs(x="Interval",y="Average Steps")
```


2: Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
avg_steps_interval[which.max(avg_steps_interval$steps), ]
```


## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.



1: Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
# Total NA
sum(is.na(activity))
```

2: Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
# average steps per day
avg_steps_day<-
  activity%>%
  group_by(date) %>%
  summarize(average_steps = mean(steps)
  ) 
avg_steps_day[is.na(avg_steps_day)]=0

act_i <- merge(activity,avg_steps_day)

```

3: Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}
# NA’s are replaced with average steps for that day
activity_imputed <- act_i %>%
 mutate(steps = ifelse(is.na(steps),average_steps , steps))

summary(activity_imputed)
```

4: Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
steps_day_1<-
  activity_imputed%>%
  group_by(date) %>%
  summarize(total_steps = sum(steps)
  ) 
head(steps_day_1)

hist(steps_day_1$total_steps, col = steps_day_1$date,main="Total Steps Per Day Imputed", xlab="Total Steps")
```


## mean
```{r}
mean(steps_day_1$total_steps)
```
## median
```{r}
median(steps_day_1$total_steps)
```



## Are there differences in activity patterns between weekdays and weekends?

```{r}
# adding weekday and weekend type
activity_imputed$daytype[weekdays(as.Date(activity_imputed$date))  %in% c('Saturday','Sunday') ] <- "weekend"
activity_imputed$daytype[!weekdays(as.Date(activity_imputed$date))  %in% c('Saturday','Sunday')] <- "weekday"

activity_imputed$daytype <- as.factor(activity_imputed$daytype)

avg_imputed_steps_interval <- aggregate(steps ~ interval + daytype, activity_imputed, mean)
qplot(interval, steps, data = avg_imputed_steps_interval, 
      geom=c("line")) + 
      facet_wrap(.~daytype,ncol=1) +      
      labs(x="Interval",y="Number of steps")
```
© 2019 GitHub, Inc.